---
title: "Untitled"
author: "Rani Yasmin Azzahra "
date: "2023-09-30"
output: html_document
---
## Packages
```{r}
library(dplyr)
library(ggplot2)
library(skimr)
library(DataExplorer)
library(RColorBrewer)
library(tidyverse)

library(VGAM)
library(rms)
```

## Input Data Train
```{r}
train <- read.csv("https://raw.githubusercontent.com/raniysmnz/DAC/main/DataTrain_Preliminary.csv", header=TRUE, sep=";")
str(train)
```
## Data Preprocessing
### Remove missing values and unusual data
The dataset is large enough to remove missing values and unusual data without imputation.
```{r}
#check missing values and summary of data
skim(train)
```

```{r}
#function to make an column is binary only
convert.bin <- function(column) {
  column[column != 0 & column != 1] <- NA
  return(column)
}
#list of binary columns
factor.b.col <- c("land", "logged_in", "root_shell", "su_attempted", "is_host_login", "is_guest_login")
#apply function to binary columns
train[factor.b.col] <- lapply(train[factor.b.col], convert.bin)
```

```{r}
#set unusual data as NA
train[train == "*"] <- NA
train[train == "99999"] <- NA
train[train == ""] <- NA
train[train == "#NA"] <- NA
```

```{r}
# remove NA data
train <- na.omit(train)
```

### Fix data type
```{r}
#check unique values for each variable
lapply(train, unique)
```


```{r}
#set data type as factor
##numeric data that contain only 2 unique number are set as factor
factor.col <- c("protocol_type", "service", "flag", "urgent", "land", "logged_in", "root_shell", "su_attempted", "is_host_login", "is_guest_login", "type_of_attack", "num_shells", "serror_rate", "srv_serror_rate", "rerror_rate", "srv_rerror_rate", "same_srv_rate", "srv_diff_host_rate", "dst_host_srv_diff_host_rate", "dst_host_srv_serror_rate", "dst_host_rerror_rate", "dst_host_srv_rerror_rate")
train[factor.col] <- lapply(train[factor.col], as.factor)

#remove "is_host_login", "diff_srv_rate", "num_outbound_cmds" column as the cleaned data just have one type of its category
train <- train %>% dplyr::select( -is_host_login, -diff_srv_rate, -num_outbound_cmds)

#set data type as numeric
num.col <- setdiff(names(train), factor.col)
train[num.col] <- lapply(train[num.col], as.numeric)
```

```{r}
str(train)
skim(train)
```

## Exploration and Visualize Data
### Distribution of Type of Attack
```{r}
ta.plot <-
  ggplot(train, aes(y=type_of_attack, fill=type_of_attack)) +
  geom_bar()+
  scale_fill_brewer(palette = "Set3")+
  labs(title="Type of Attack Distribution")+
  theme(plot.title = element_text(hjust = 0.5))
ta.plot
```

### Distribution of Protocol Type
```{r}
protocoltype.plot <-
  ggplot(train, aes(y=protocol_type, fill=protocol_type)) +
  geom_bar()+
  scale_fill_brewer(palette = "Set3")+
  labs(title="Protocol Type Distribution")+
  theme(plot.title = element_text(hjust = 0.5))
protocoltype.plot
```
### Distribution of Flag Type
```{r}
flag.plot <-
  ggplot(train, aes(y=flag, fill=flag)) +
  geom_bar()+
  scale_fill_brewer(palette = "Set3")+
  labs(title="Flag Type Distribution")+
  theme(plot.title = element_text(hjust = 0.5))
flag.plot
```
### Density Plot for Numeric Variables
```{r}
int.density <-
  plot_density(train[num.col])
```

## Ordinal Logistic Regression
### Modelling
```{r}
library(nnet)
# Since we are going to use Academic as the reference group, we need relevel the group.
train$type_of_attack <- relevel(as.factor(train$type_of_attack), ref = "normal")

# Run a "only intercept" model
incept.multilog.reg <- multinom(type_of_attack ~ ., data = train)

# 
multilog.reg <- multinom(type_of_attack ~ ., data = train, model=TRUE)
#summary(multilog.reg)
```














